### 2021-04-20 06_presentations/210421_cqm/index.Rmd
updating presentation for cqm

---
title: "‘Dissecting the causal question underlying the association between cancer and dementia"
subtitle: ""  
author: 
  - "L. Paloma Rojas-Saunero"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      highlightLanguage: ["r", "css"]
      countIncrementalSlides: FALSE
      ratio: "16:9"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE, }
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(collapse = TRUE,
                      fig.retina = 3,
                      warning = FALSE,
                      message = FALSE, cache = TRUE)
library(WeightIt)
library(cobalt)
library(splines)

source(here::here("02_R", "05_plots.R"))
source(here::here("02_R", "05b_plot_causes_death.R"))
table_results <- rio::import(here::here("02_R", "table_results.csv"))
table_results_bounds <- rio::import(here::here("02_R", "table_results_bounds.csv"))

```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#011A5E",
  header_font_google = google_font("Fira Sans", "500"),
  table_row_even_background_color = "#FFFFFF",
  table_border_color = "#011A5E",
  text_font_google   = google_font("Lato", "400", "400i"),
  text_font_size = "25px"
)
```

background-image: url(./figs/erasmus.png)
background-position: 5% 95%
background-size: 25%, 15%
class: center, middle

## Dissecting the causal question underlying the association between cancer and dementia

<br> <br>

.right[L. Paloma Rojas-Saunero, MD

**Epidemiology department**

**Causal inference and Neuroepi group**]

---

## Introduction

.pull-left[

Observational studies have found a protective effect of cancer diagnosis in the risk of Alzheimer´s disease or related dementias (ADRD).

These studies lead to discussions of repurposing or augmenting current cancer chemotherapeutics for ADRD
]

--

.pull-right[ 
![](figs/forest_plot.PNG)
]

.footnote[Ospina-Romero M. JAMA Network Open. 2020]

---

## Cancer diagnosis as an exposure

.pull-left[

- Molecular mechanisms

- Cancer treatments

- Shared factors

- Unmeasured confounding, measurement error and survival bias may explain this association

]

.pull-right[

![](figs/causal_risks.PNG)
]

.footnote[
L. Frain et al. Alzheimer’s & Dementia. 2017.
]

---
class: middle, center

## How can we conclude on causal pathways if we haven't measured them?

---

## Back to basics

.pull-left[
![](figs/pin_knockout.PNG) 
]

.pull-right[

Driver J. et al. Biochimica et Biophysica Acta. 2015.

![](figs/tment_pin1.PNG)

]

---

## Aims

---

### Overview of cancer diagnosis as a proxy for Pin1 enzime

.pull-left[
If there was a drug that targets Pin1 $(D)$ we would perform an RCT to assess the effect on dementia $(Y_{t})$.

We can imagine a setting where middle-aged participants with signs of cognitive decline take this drug and are followed over time.
]

--

<br><br>

.pull-right[
```{tikz, fig.ext = 'png', echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm, auto,>=latex, scale = 0.5]
\node (p) {$D$};
\node [right of = p] (y) {$Y_{t}$};
\draw[->] (p) -- (y);
\end{tikzpicture}
```
]

---
class: middle

.pull-left[
- The drug is not available yet, but we have better chances of having Pin1 expression as a biomarker in a closer future.

- This biomarker could be measured in stored blood samples in a cohort study

- We can set a threshold for _Pin1 over-expression_ = $P$

- Since this is within an observational setting we will have to address confounding $(L)$

]

.pull-right[
```{tikz, fig.ext = 'png', echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm, auto,>=latex, scale = 0.5]
\node (p) {$P$};
\node [below of = p, yshift = 1cm, xshift = -1cm](l) {$L$};
\node [right of = p] (y) {$Y_{t}$};
\draw[->] (l) -- (p);
\draw[->, violet] (p) -- (y);
\draw[->] (l) -- (y);
\end{tikzpicture}
```
]

---

class: middle

.pull-left[

- Pin1 over-expression stimulates tumor growth

- Tumor growth leads to cancer diagnosis

- Cancer diagnosis $(P^*)$ represents a proxy for $P$

- We are assuming that the captured effect of $P^*$ and $Y_{t}$ is only through $P$
]

.pull-right[
```{tikz, fig.ext = 'png', echo = FALSE, fig.cap = ""}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm, auto,>=latex, scale = 0.5]
\node (p) {$P$};
\node [below of = p, yshift = 1cm, xshift = -1cm](l) {$L$};
\node [above of = p, yshift = -1cm](pstar) {$P^*$};
\node [right of = p] (y) {$Y_{t}$};
\draw[->, violet] (p) -- (pstar);
\draw[->, violet] (p) -- (y);
\draw[->] (l) -- (p);
\draw[->] (l) -- (y);

\end{tikzpicture}
```
]

???

Not everyone who had Pin1 over-expression will be diagnosed with cancer by late-midlife, but in late-life.

---
class:middle

.pull-left[
Some people will die before having a cancer diagnosis

]

.pull-right[
```{tikz, fig.ext = 'png', echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm, auto,>=latex, scale = 0.5]
\node (p) {$P_0$};
\node [right of = p] (d) {$D_{k} = 0$};
\node [right of = d] (pstar) {$P^*_{k+1}$};
\node [below of = p, yshift = 1cm, xshift = -1cm](l) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm](c) {$C_2$};
\draw[->] (p) -- (d);
\draw[->, violet] (p) to [out=45,in=135, looseness=0.4] (pstar);
\draw[->] (d) -- (pstar);
\draw[->] (l) -- (p);
\draw[->] (l) -- (d);
\draw[->] (c) -- (pstar);
\draw[->] (c) -- (d);
\end{tikzpicture}
```
]

---
class: middle, center

## Overview of the causal question and assumptions related to death as a competing event of ADRD

---
class: middle

.pull-left[

]

.pull-right[
```{tikz, fig.ext = 'png', echo = FALSE, fig.cap = ""}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm, auto,>=latex, scale = 0.5]
\node (p) {$P$};
\node [below of = p, yshift = 1cm, xshift = -1cm](l) {$L$};
\node [above of = p, yshift = -1cm](pstar) {$P^*$};
\node [right of = p] (d) {$D_{t}$};
\node [right of = d] (y) {$Y_{t+1}$};
\draw[->, violet] (p) -- (pstar);
\draw[->] (p) -- (d);
\draw[->] (d) -- (y);
\draw[->] (l) -- (p);
\draw[->] (l) -- (y);
\draw[->, violet] (p) to [out=45,in=135, looseness=0.4] (y);
\end{tikzpicture}
```
]

---
class: middle

.pull-left[]

.pull-right[
```{tikz, fig.ext = 'png', echo = FALSE, fig.cap = ""}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm, auto,>=latex, scale = 0.5]
\node (p) {$P$};
\node [below of = p, yshift = 1cm, xshift = -1cm](l) {$L$};
\node [above of = p, yshift = -1cm](pstar) {$P^*$};
\node [right of = p] (d) {$D_{t}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = d, yshift = 1cm, xshift = +1cm](c) {$C$};
\draw[->, violet] (p) -- (pstar);
\draw[->] (pstar) -- (d);
\draw[->] (p) -- (d);
\draw[->] (d) -- (y);
\draw[->] (l) -- (p);
\draw[->] (l) -- (y);
\draw[->] (c) -- (d);
\draw[->] (c) -- (y);
\draw[->, violet] (p) to [out=45,in=135, looseness=0.4] (y);
\end{tikzpicture}
```

]

---

## Application to the Rotterdam Study

.pull-left[

- Participants between 60 and 70 years old at study entry
- Without history of cancer diagnosis and dementia
- Data on cancer diagnosis and dementia diagnosis is measured through follow-up and linkage with national registries.
- Date and cause of death was collected on a weekly basis via municipal population registries

]

.pull-right[
```{r, echo = FALSE}
plot_distribution +
    theme_xaringan()
```
]

---

### Methods

---

### Results

```{r, echo = FALSE}
kableExtra::kable(table_results, booktabs = T) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position"))
```

---

### Extreme scenarios of independency

```{r, echo = FALSE, fig.showtext=TRUE, fig.align='center', fig.width = 14}
table_results_bounds %>% 
  select(-5, -6) %>%
  pivot_longer(cols = 3:4,
               names_to = "cancer",
               values_to = "risk") %>%
  pivot_wider(names_from = c(model),
              values_from = risk) %>%
  ggplot(aes(
    y = Proxy,
    x = IPTW,
    xmin = `Lower bound`,
    xmax = `Upper bound`
  )) +
  geom_vline(xintercept = 1,
             color = "#E69F00",
             size = 1.2) +
  geom_pointrange(shape = 15,
                  color = "#0072B2") +
  geom_errorbar(width = 0.2, color = "#0072B2") + facet_wrap(. ~ cancer) +
  theme_xaringan()
```

---

```{r, echo = FALSE}
kableExtra::kable(table_results_bounds, booktabs = T) %>%
  kableExtra::kable_styling(latex_options = c("striped", "hold_position")) 
```

---

### Causes of death

```{r, echo = FALSE, fig.align='center', fig.width = 14}
causes_death + 
      theme_xaringan(text_font_size = 20,
                       title_font_size = 20) +
     theme(legend.position = "right")

```