### 2021-04-27 03_figs/tikz364831c2bf8.tex
ser presentation updates

\documentclass[tikz]{standalone}
\usepackage{amsmath}
\usetikzlibrary{matrix}
%% EXTRA_TIKZ_PREAMBLE_CODE %%
\begin{document}
%% TIKZ_CODE %%
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm, auto,>=latex, scale = 0.5]
\node (p) {$P$};
\node [below of = p, yshift = 1cm, xshift = -1cm, gray!50](l) {$L$};
\node [right of = p, gray!50] (dp) {$D_t$};
\node [right of = dp, gray!50] (pstar) {$P*_t$};
\node [right of = pstar, gray!50] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [above of = dp, yshift = -1cm, xshift = +1cm, gray!50](c) {$C_1$};
\node [above of = d, yshift = -1cm, xshift = +1cm, gray!50](cd) {$C_2$};
\draw[->, gray!50] (p) -- (dp);
\draw[->, gray!50] (dp) -- (pstar);
\draw[->, gray!50] (p) to [out=-45,in=-135, looseness=0.3] (pstar);
\draw[->, gray!50] (pstar) -- (d);
\draw[->, gray!50] (p) to [out=-45,in=-135, looseness=0.3] (d);
\draw[->] (p) to [out=-45,in=-135, looseness=0.3] (y);
\draw[->, gray!50] (d) -- (y);
\draw[->, gray!50] (l) -- (p);
\draw[->, gray!50] (l) to [out=0,in=270, looseness=0.4] (y);
\draw[->, gray!50] (c) -- (dp);
\draw[->, gray!50] (c) -- (pstar);
\draw[->, v] (cd) -- (d);
\draw[->, gray!50] (cd) -- (y);
\end{tikzpicture}
\end{document}