### 2021-04-27 06_presentations/ser2021/index.Rmd

---
title: "Toward a clearer understanding of the inverse association between cancer and dementia"
subtitle: ""  
author: 
  - "L. Paloma Rojas-Saunero"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      highlightLanguage: ["r", "css"]
      countIncrementalSlides: FALSE
      ratio: "16:9"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE, fig.showtext = TRUE, cache = TRUE)

library(splines)
library(knitr)
library(kableExtra)

source(here::here("02_R", "04_direct_effect_simplified.R"))
source(here::here("02_R", "05_plots.R"))
source(here::here("02_R", "05b_plot_causes_death.R"))
table_results <- rio::import(here::here("02_R", "table_results.csv"))
table_results_bounds <- rio::import(here::here("02_R", "table_results_bounds.csv"))

```

```{r xaringan-themer, include=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#011A5E",
  header_font_google = google_font("Fira Sans", "500"),
  table_row_even_background_color = "#FFFFFF",
  table_border_color = "#011A5E",
  text_font_google   = google_font("Lato", "400", "400i"),
  text_font_size = "30px"
)
```

background-image: url(./figs/ser.jpg), url(./figs/erasmus.png)
background-position: 5% 95%, 95% 5%
background-size: 25%, 15%
class: center, middle

## Towards a clearer understanding of the inverse association between cancer and dementia

<br> <br> <br>

.right[L. Paloma Rojas-Saunero, MD

**Epidemiology department**]

---
## Introduction

<br><br>

- Observational studies have found an inverse association between cancer and the risk of Alzheimer´s disease or related dementias (ADRD).

--

- To understand the potential sources of bias we first need to define the causal question.

--

- Refining the underlying question behind cancer diagnosis as the exposure unveils some of these concerns.

---

### Pin1-targeting drug

```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_0$};
\node [above of = p, yshift = -1cm, xshift = 1cm, gray!30](l) {$L$};
\node [right of = p, gray!30] (dp) {$D_t$};
\node [right of = dp, gray!30] (pstar) {$P^*_t$};
\node [right of = pstar, gray!30] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = dp, yshift = 1cm, xshift = 1cm, gray!30](c) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm, gray!30](cd) {$C_2$};
\draw[->, gray!30] (p) -- (dp);
\draw[->, gray!30] (l) -- (p);
\draw[->, gray!30] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->, gray!30] (dp) -- (pstar);
\draw[->, gray!30] (p) to [out=+45,in=+135, looseness=0.3] (pstar);
\draw[->, gray!30] (pstar) -- (d);
\draw[->, gray!30] (p) to [out=+45,in=+135, looseness=0.3] (d);
\draw[->, gray!30] (d) -- (y);
\draw[->, gray!30] (c) -- (dp);
\draw[->, gray!30] (c) -- (pstar);
\draw[->, gray!30] (cd) -- (d);
\draw[->, gray!30] (cd) -- (y);
\end{tikzpicture}
```

$P_0$ = Pin1 over-expression; $Y_{t+1}$ = Dementia diagnosis

???
In hypothetical setting, if Pin1-targeting drug was developed, the best way to understand its effect on dementia risk would be to do a randomized trial in which we randomize eligible participants in late midlife to receive this drug or not, and closely monitor ADRD over follow-up.

---
### Pin1 expression as biomarker (Problem 1: _Confounding_)

```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_0$};
\node [above of = p, yshift = -1cm, xshift = 1cm](l) {$L$};
\node [right of = p, gray!30] (dp) {$D_t$};
\node [right of = dp, gray!30] (pstar) {$P^*_t$};
\node [right of = pstar, gray!30] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = dp, yshift = 1cm, xshift = 1cm, gray!30](c) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm, gray!30](cd) {$C_2$};
\draw[->, gray!30] (p) -- (dp);
\draw[->] (l) -- (p);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->, semithick] (p) to [out=45,in=135, looseness=0.3] (y);
\draw[->, gray!30] (dp) -- (pstar);
\draw[->, gray!30] (p) to [out=45,in=135, looseness=0.3] (pstar);
\draw[->, gray!30] (pstar) -- (d);
\draw[->, gray!30] (p) to [out=45,in=135, looseness=0.3] (d);
\draw[->, gray!30] (d) -- (y);
\draw[->, gray!30] (c) -- (dp);
\draw[->, gray!30] (c) -- (pstar);
\draw[->, gray!30] (cd) -- (d);
\draw[->, gray!30] (cd) -- (y);
\end{tikzpicture}
```

$P_0$ = Pin1 over-expression; $Y_{t+1}$ = Dementia diagnosis; $L$ = Shared causes of $P_0$ and $Y_{t+1}$

???
We might be closer to developing biomarker for Pin1 expression in the future. So we can imagine a situation in which Pin1 exposure is measured in a population based cohort study, using stored blood samples.
Since this would be a observational setting, Pin1 expression and ADRD at time may share causes such as smoking, and that assessing the causal relationship requires adjusting for these confounders.

---
### Cancer diagnosis as the proxy of Pin1 biomarker (Problem 2: _Time zero_)

```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_0$};
\node [above of = p, yshift = -1cm, xshift = 1cm, gray!30](l) {$L$};
\node [right of = p, gray!30] (dp) {$D_t$};
\node [right of = dp] (pstar) {$P^*_t$};
\node [right of = pstar, gray!30] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = dp, yshift = 1cm, xshift = 1cm, gray!30](c) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm, gray!30](cd) {$C_2$};
\draw[->, gray!30] (p) -- (dp);
\draw[->, gray!30] (l) -- (p);
\draw[->, gray!30] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->, gray!30] (dp) -- (pstar);
\draw[->] (p) to [out=45,in=135, looseness=0.3] (pstar);
\draw[->, gray!30] (pstar) -- (d);
\draw[->, gray!30] (p) to [out=+45,in=+135, looseness=0.3] (d);
\draw[->, gray!30] (d) -- (y);
\draw[->, gray!30] (c) -- (dp);
\draw[->, gray!30] (c) -- (pstar);
\draw[->, gray!30] (cd) -- (d);
\draw[->, gray!30] (cd) -- (y);
\end{tikzpicture}
```

$P_0$ = Pin1 over-expression; $Y_{t+1}$ = Dementia diagnosis; $P^*_t$ = Cancer diagnosis

???
At this moment this biomarker is not available, but we know that Pin1 over-expression stimulates tumor growth and for this reason we use cancer diagnosis as the proxy for Pin1.

---
### Death prior to cancer diagnosis (Problem 3: _Survival bias_)

```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_0$};
\node [above of = p, yshift = -1cm, xshift = 1cm, gray!30](l) {$L$};
\node [right of = p] (dp) {$D_t$};
\node [right of = dp] (pstar) {$P^*_t$};
\node [right of = pstar, gray!30] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = dp, yshift = 1cm, xshift = 1cm, gray!30](c) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm, gray!30](cd) {$C_2$};
\draw[->] (p) -- (dp);
\draw[->, gray!30] (l) -- (p);
\draw[->, gray!30] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->] (dp) -- (pstar);
\draw[->] (p) to [out=45,in=135, looseness=0.3] (pstar);
\draw[->, gray!30] (pstar) -- (d);
\draw[->, gray!30] (p) to [out=+45,in=+135, looseness=0.3] (d);
\draw[->, gray!30] (d) -- (y);
\draw[->, gray!30] (c) -- (dp);
\draw[->, gray!30] (c) -- (pstar);
\draw[->, gray!30] (cd) -- (d);
\draw[->, gray!30] (cd) -- (y);
\end{tikzpicture}
```

$P_0$ = Pin1 over-expression; $Y_{t+1}$ = Dementia diagnosis; $P^*_t$ = Cancer diagnosis; $D_t$ = Death prior to cancer diagnosis
???

---
### Death prior to cancer diagnosis (Problem 3: _Survival bias_)

```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_0$};
\node [above of = p, yshift = -1cm, xshift = 1cm, gray!30](l) {$L$};
\node [right of = p] (dp) {$D_t$};
\node [right of = dp] (pstar) {$P^*_t$};
\node [right of = pstar, gray!30] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = dp, yshift = 1cm, xshift = 1cm](c) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm, gray!30](cd) {$C_2$};
\draw[->] (p) -- (dp);
\draw[->, gray!30] (l) -- (p);
\draw[->, gray!30] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->] (dp) -- (pstar);
\draw[->] (p) to [out=45,in=135, looseness=0.3] (pstar);
\draw[->, gray!30] (pstar) -- (d);
\draw[->, gray!30] (p) to [out=+45,in=+135, looseness=0.3] (d);
\draw[->, gray!30] (d) -- (y);
\draw[->] (c) -- (dp);
\draw[->] (c) -- (pstar);
\draw[->, gray!30] (cd) -- (d);
\draw[->, gray!30] (cd) -- (y);
\end{tikzpicture}
```

$P_0$ = Pin1 over-expression; $Y_{t+1}$ = Dementia diagnosis; $P^*_t$ = Cancer diagnosis; $D_t$ = Death prior to cancer diagnosis; $C_1$ = Shared causes of $D_t$ and $P^*$
???

---
### Death prior to dementia diagnosis (Problem 3: _Survival bias_)

```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_0$};
\node [above of = p, yshift = -1cm, xshift = 1cm, gray!30](l) {$L$};
\node [right of = p, gray!30, gray!30] (dp) {$D_t$};
\node [right of = dp, gray!30] (pstar) {$P^*_t$};
\node [right of = pstar] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = dp, yshift = 1cm, xshift = 1cm, gray!30](c) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm, gray!30](cd) {$C_2$};
\draw[->, gray!30] (p) -- (dp);
\draw[->, gray!30] (l) -- (p);
\draw[->, gray!30] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->, gray!30] (dp) -- (pstar);
\draw[->, gray!30] (p) to [out=45,in=135, looseness=0.3] (pstar);
\draw[->, gray!30] (pstar) -- (d);
\draw[->] (p) to [out=+45,in=+135, looseness=0.3] (d);
\draw[->] (d) -- (y);
\draw[->, gray!30] (c) -- (dp);
\draw[->, gray!30] (c) -- (pstar);
\draw[->, gray!30] (cd) -- (d);
\draw[->, gray!30] (cd) -- (y);
\end{tikzpicture}
```

$P_0$ = Pin1 over-expression; $Y_{t+1}$ = Dementia diagnosis; $D_{t+1}$ = Death prior to dementia diagnosis

---
### Death prior to dementia diagnosis (Problem 3: _Survival bias_)

```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_0$};
\node [above of = p, yshift = -1cm, xshift = 1cm, gray!30](l) {$L$};
\node [right of = p, gray!30, gray!30] (dp) {$D_t$};
\node [right of = dp, gray!30] (pstar) {$P^*_t$};
\node [right of = pstar] (d) {$D_{t+1}$};
\node [right of = d] (y) {$Y_{t+1}$};
\node [below of = dp, yshift = 1cm, xshift = 1cm, gray!30](c) {$C_1$};
\node [below of = d, yshift = 1cm, xshift = 1cm](cd) {$C_2$};
\draw[->, gray!30] (p) -- (dp);
\draw[->, gray!30] (l) -- (p);
\draw[->, gray!30] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->, gray!30] (dp) -- (pstar);
\draw[->, gray!30] (p) to [out=45,in=135, looseness=0.3] (pstar);
\draw[->, gray!30] (pstar) -- (d);
\draw[->] (p) to [out=+45,in=+135, looseness=0.3] (d);
\draw[->] (d) -- (y);
\draw[->, gray!30] (c) -- (dp);
\draw[->, gray!30] (c) -- (pstar);
\draw[->] (cd) -- (d);
\draw[->] (cd) -- (y);
\end{tikzpicture}
```

$P_0$ = Pin1 over-expression; $Y_{t+1}$ = Dementia diagnosis; $D_{t+1}$ = Death prior to dementia diagnosis; $C_2$ as shared causes of $D_{t+1}$ and $Y_{t+1}$

---
## Application to the Rotterdam Study

- Participants between 60 and 70 years old at study entry, free of cancer and dementia diagnosis at baseline

--

- Data on cancer diagnosis, dementia diagnosis and death was collected from national registries and electronical clinical records.

--

- Cancer as _"time-fixed ever vs. never"_ vs. cancer as a _time-varying proxy_

---
## Application to the Rotterdam Study

- Inverse probability treatment weights to address confounding (IPTW)

  + *Age, sex, Apoe4, education*

--

- Inverse probability censoring weights for death (IPCW)

  + *Age, sex, Apoe4, education and time-varying covariates: smoking, systolic blood pressure, BMI and incident cancer, heart disease, stroke and diabetes*

--

- Risk of dementia is estimated with a weighted Kaplan-Meier estimator.

---
### Peterson upper and lower bounds

- **1) Lower Bound**: Total independence, that refers to an scenario were those who died would never develop dementia (Death as competing event)
--

- **2) Upper Bound**: Complete dependency, that refers to an scenario where those who died would have dementia prior to death (Composite outcome, death + dementia)

---
### Results

.center[
```{r, echo = FALSE, fig.retina = 3}
plot_distribution +
    theme_xaringan()
```

In the cancer arm 6% had dementia and 63% died, in the free of cancer arm 14% had dementia and 23% died.

]

---
### Cancer as time-fixed ever vs. never proxy

.pull-left[
```{r}
plot_1b
```

]

--

.pull-right[
```{r}
plot_1c
```

]

---

### Cancer diagnosis as time-varying proxy

.pull-left[
```{r}
plot_2b
```

]

--

.pull-right[
```{r}
plot_2c
```

]
---
### Results

<br>

.center[
```{r, echo = FALSE}
tb <- table_results %>%
  filter(model != "Unadjusted") %>%
  select(-c(3, 4, 5)) %>%
  rename(RR = rr, `HR (95%CI)` = hr) %>% 
  filter(Proxy != "Time to cancer")

tb %>% kbl() %>%
kable_styling(bootstrap_options = c("striped", "hover"))
```
]
---

### Extreme scenarios of independency between death and dementia

.pull-left[
```{r, echo = FALSE}
table_results_bounds %>%
  select(-3, -4, - 6) %>%
    pivot_wider(names_from = c(model),
              values_from = rd) %>%
  ggplot(aes(
    y = Proxy,
    x = `IPTW + IPCW`,
    xmin = `Lower bound`,
    xmax = `Upper bound`
  )) +
  geom_vline(xintercept = 0,
             color = "#E69F00",
             size = 1.2) +
  geom_pointrange(shape = 15,
                  color = "#0072B2") +
  geom_errorbar(width = 0.2, color = "#0072B2") +
  labs(x = "Risk difference",
       y = NULL) +
  theme_xaringan()
```
]

--

.pull-right[
```{r, echo = FALSE}
table_results_bounds %>%
  select(-3, -4, - 5) %>%
    pivot_wider(names_from = c(model),
              values_from = rr) %>%
  ggplot(aes(
    y = Proxy,
    x = `IPTW + IPCW`,
    xmin = `Lower bound`,
    xmax = `Upper bound`
  )) +
  geom_vline(xintercept = 1,
             color = "#E69F00",
             size = 1.2) +
  geom_pointrange(shape = 15,
                  color = "#0072B2") +
  geom_errorbar(width = 0.2, color = "#0072B2") +
  labs(x = "Risk Ratio",
       y = NULL) +
   scale_x_continuous(
    trans = "log10",
    # breaks = seq(0.7, 1.2, by = 0.1)
  ) +
  theme_xaringan()
```
]

---
### Conclusions

- The process of connecting our real question to the data help us identify the structure that will later represent the analytic decisions and assumptions.

--

- Cancer as the proxy for a molecular mechanism is tricky...
--

- When death is a competing event, prior to choosing the estimator we need to define what is the estimand of interest and be explicit about the assumptions and interpretations.

---