### 2021-07-13 04_reports/dags.Rmd
its been a while (new edits to Sonja's edits)

---
title: "Dags"
output: word_document
---
## Current DAG
```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_{t-1}$};
\node [above of = p, yshift = -0.8cm, xshift = -1cm](l) {$L$};
\node [below of = p, yshift = 1cm] (pstar) {$P^*_{t-1}$};
\node [right of = p] (dp) {$D_t$};
\node [right of = dp] (p2) {$P_t$};
\node [below of = p2, yshift = 1cm] (pstar2) {$P^*_t$};
\node [right of = p2] (dp2) {$D_{t+1}$};
\node [right of = dp2] (y) {$Y_{t+1}$};
\node [below of = dp2, yshift = 1cm, xshift = +1cm](cd) {$C$};
\draw[->] (p) to (pstar);
\draw[->] (l) -- (p);
\draw[->] (l) to (dp);
\draw[->] (pstar) -- (dp);
\draw[->] (dp) to [out=-45,in=-135, looseness=0.3] (dp2);
\draw[->] (dp) -- (p2);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->] (p2) to (pstar2);
\draw[->] (dp2) -- (y);
\draw[->] (cd) -- (dp2);
\draw[->] (cd) -- (y);
\draw[->] (l) -- (p2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (dp2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->] (pstar2) -- (dp2);
\draw[->, semithick] (p2) to [out=+45,in=+135, looseness=0.3] (y);
\end{tikzpicture}
```

---
## Changing time and L
```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_t$};
\node [above of = p, yshift = -0.8cm, xshift = -1cm](l) {$L_1$};
\node [below of = p, yshift = 1cm] (pstar) {$P^*_t$};
\node [right of = p] (dp) {$D_{t+1}$};
\node [right of = dp] (p2) {$P_{t+1}$};
\node [below of = p2, yshift = 1cm] (pstar2) {$P^*_{t+1}$};
\node [right of = p2] (dp2) {$D_{t+k}$};
\node [right of = dp2] (y) {$Y_{t+k}$};
\node [below of = dp2, yshift = 1cm, xshift = +1cm](cd) {$L_2$};
\draw[->] (p) to (pstar);
\draw[->] (l) -- (p);
\draw[->] (l) to (dp);
\draw[->] (pstar) -- (dp);
\draw[->] (dp) to [out=-45,in=-135, looseness=0.3] (dp2);
\draw[->] (dp) -- (p2);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->] (p2) to (pstar2);
\draw[->] (dp2) -- (y);
\draw[->] (cd) -- (dp2);
\draw[->] (cd) -- (y);
\draw[->] (l) -- (p2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (dp2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->] (pstar2) -- (dp2);
\draw[->, semithick] (p2) to [out=+45,in=+135, looseness=0.3] (y);
\end{tikzpicture}
```

---
## Changing time, L and proxy
```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_t$};
\node [above of = p, yshift = -0.8cm, xshift = -1cm](l) {$L_1$};
\node [below of = p, yshift = 1cm] (pstar) {$C_t$};
\node [right of = p] (dp) {$D_{t+1}$};
\node [right of = dp] (p2) {$P_{t+1}$};
\node [below of = p2, yshift = 1cm] (pstar2) {$C_{t+1}$};
\node [right of = p2] (dp2) {$D_{t+k}$};
\node [right of = dp2] (y) {$Y_{t+k}$};
\node [below of = dp2, yshift = 1cm, xshift = +1cm](cd) {$L_2$};
\draw[->] (p) to (pstar);
\draw[->] (l) -- (p);
\draw[->] (l) to (dp);
\draw[->] (pstar) -- (dp);
\draw[->] (dp) to [out=-45,in=-135, looseness=0.3] (dp2);
\draw[->] (dp) -- (p2);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->] (p2) to (pstar2);
\draw[->] (dp2) -- (y);
\draw[->] (cd) -- (dp2);
\draw[->] (cd) -- (y);
\draw[->] (l) -- (p2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (dp2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->] (pstar2) -- (dp2);
\draw[->, semithick] (p2) to [out=+45,in=+135, looseness=0.3] (y);
\end{tikzpicture}
```

---

## Changing time, L and proxy, plus L to C
```{tikz, echo = FALSE}
\usetikzlibrary{arrows}
\begin{tikzpicture}[node distance=2cm]
\node (p) {$P_t$};
\node [above of = p, yshift = -0.8cm, xshift = -1cm](l) {$L_1$};
\node [below of = p, yshift = 1cm] (pstar) {$C_t$};
\node [right of = p] (dp) {$D_{t+1}$};
\node [right of = dp] (p2) {$P_{t+1}$};
\node [below of = p2, yshift = 1cm] (pstar2) {$C_{t+1}$};
\node [right of = p2] (dp2) {$D_{t+k}$};
\node [right of = dp2] (y) {$Y_{t+k}$};
\node [below of = dp2, yshift = 1cm, xshift = +1cm](cd) {$L_2$};
\draw[->] (p) to (pstar);
\draw[->] (l) -- (p);
\draw[->] (l) -- (pstar);
\draw[->] (l) to (dp);
\draw[->] (pstar) -- (dp);
\draw[->] (dp) to [out=-45,in=-135, looseness=0.3] (dp2);
\draw[->] (dp) -- (p2);
\draw[->, semithick] (p) to [out=+45,in=+135, looseness=0.3] (y);
\draw[->] (p2) to (pstar2);
\draw[->] (dp2) -- (y);
\draw[->] (cd) -- (dp2);
\draw[->] (cd) -- (y);
\draw[->] (l) -- (p2);
\draw[->] (l) to [out=-45,in=-180, looseness=0.4] (pstar2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (dp2);
\draw[->] (l) to [out=0,in=90, looseness=0.4] (y);
\draw[->] (pstar2) -- (dp2);
\draw[->, semithick] (p2) to [out=+45,in=+135, looseness=0.3] (y);
\end{tikzpicture}
```