### 2021-10-26 04_reports/supplemental.Rmd
version to submit to Epidemiology

---
title: "supplemental"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Supplementary material

### 1. Modeling description

### 2. Figure 1. Distribution of participants under each health status, by age over follow-up

```{r}
knitr::include_graphics(here::here("03_figs", "plot_distribution.tiff"))
```

### 3. Figure 2. Causes of death for participants who died prior to dementia diagnosis

```{r}
knitr::include_graphics(here::here("03_figs", "causes_death.tiff"))
```